---
title: "Covid91 vaccine study FinalFall2022"
author: "Anthony Benitez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalFall2022.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
M = filter(FinalData, sex =="M")
F = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Methods
We have to change the data in final data and final dat to factor variables then sort out the ones who were treated with the placebo and the vaccine and the others who are infected with covid with the same way and see what we get from it and see what are the diffrence between males, females, druggies, and gays.



## Males

## Graphical


```{r}
basic <- M %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
basicC <- ggplot(basic,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
basicCC <- ggplot(basic,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This bar chart shows that the men who got covid and got infected with the placebo are about 60% and the ones who got the vaccine got infected by about 30%. 

## Numerical
```{r}
table1 <- xtabs(~infected + treatment, data=M)
rowPerc(table1)
colPerc(table1)
```

So the men who took the placebo and got infected is 68.44% and the ones who took the vaccine only 31.56% got infected.

## Inferential 
```{r}
chisq.test(table1)
chisqtestGC(table1)
Mfish <- fisher.test(table1)
Mfish
```

For men the 95 percent confidence interval is 1.732 for the placebo and 2.752 for the vaccine the odds ratio is 2.178 the p value = 2.364e-12.
```{r}
effect <- Mfish$p.value/(Mfish$p.value +1)
effectivenessM = Mfish$estimate/(Mfish$estimate +1)*100
print(effectivenessM)
```
The odds ratio is 68.534.

```{r}
low=(Mfish$conf.int[1]/(Mfish$conf.int[1] +1))*100
hi=(Mfish$conf.int[2]/(Mfish$conf.int[2] +1))*100
options(digits=2)
```

The hi is 73.347 and the low is 63.401

## Females

## Graphical

```{r}
basic <- F %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
basicC <- ggplot(basic,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
basicCC <- ggplot(basic,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The females who had covid and took the vaccine about 58% of them got sick and the ones who took the placebo about 42% got covid and the ones who didn't get sick the ones who got the placebo it is about 50% and about 49% took the vaccine.

## Numerical
```{r}
table1 <- xtabs(~infected + treatment, data=F)
rowPerc(table1)
colPerc(table1)
```

The females who had covid only 42% had the placebo and the ones that had the vaccine only 57% of them got sick.

## Inferential
```{r}
chisq.test(table1)
chisqtestGC(table1)
Ffish <- fisher.test(table1)
Ffish
```

The p value = 0.005, the 95 percent confidence intervals are 0.60 and 0.92 the odds ratio 0.74. 
```{r}
effect <- Ffish$p.value/(Ffish$p.value +1)
effectivenessF = Ffish$estimate/(Ffish$estimate +1)*100
print(effectivenessF)
```

the odds ratio is 43

```{r}
low=(Ffish$conf.int[1]/(Ffish$conf.int[1] +1))*100
hi=(Ffish$conf.int[2]/(Ffish$conf.int[2] +1))*100
options(digits=2)
```

The hi variable is 47.858 and the low is 37.337

## LGBTQ

## Graphical 
```{r}
basic <- gay %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
basicC <- ggplot(basic,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
basicCC <- ggplot(basic,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The ones who got covid and took the placebo and got infected was about 10% and the ones who took the vaccine and got infected were about 78%. The ones who took the placebo and didn't get infected there was only 49% and the ones who took the vaccine and didn't get infected are about 50%.

## Numerical
```{r}
table1 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```

The people who got covid and took the placebo only 13% got sick and the for the ones who had the vaccine only 87% got sick.

## Inferential 
```{r}
chisq.test(table1)
chisqtestGC(table1)
gayfish <- fisher.test(table1)
gayfish
```

The p-value that i got was 1e-08 the 95 percent confidence intervals are 0.066 and 0.344 the odds ratio is 0.16.

```{r}
effect <- gayfish$p.value/(gayfish$p.value +1)
effectivenessgay = gayfish$estimate/(gayfish$estimate +1)*100
print(effectivenessgay)
```

The odds of this truly working is 14.

```{r}
low=(gayfish$conf.int[1]/(gayfish$conf.int[1] +1))*100
hi=(gayfish$conf.int[2]/(gayfish$conf.int[2] +1))*100
options(digits=2)
```

The highest is 25.613 and the lowest is 6.201

## Druggies

## Graphical

## Numerical

## Inferential 

# Overall Results and Conclusions



