---
title: "Covid91 vaccine study FinalFall2022"
author: "Anthony Benitez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalFall2022.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
M = filter(FinalData, sex =="M")
F = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Methods
We have to change the data in final data and final dat to factor variables then sort out the ones who were treated with the placebo and the vaccine and the others who are infected with covid with the same way and see what we get from it and see what are the diffrence between males, females, druggies, and gays.



## Males

## Graphical


```{r}
basic <- M %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
basicC <- ggplot(basic,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
basicCC <- ggplot(basic,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This bar chart shows that the men who got covid and got infected with the placebo are about 60% and the ones who got the vaccine got infected by about 30%. 

## Numerical
```{r}
table1 <- xtabs(~infected + treatment, data=M)
rowPerc(table1)
colPerc(table1)
```

So the men who took the placebo and got infected is 68.44% and the ones who took the vaccine only 31.56% got infected.

## Inferential 
```{r}
chisq.test(table1)
chisqtestGC(table1)
Mfish <- fisher.test(table1)
Mfish
```

For men the 95 percent confidence interval is 1.732 for the placebo and 2.752 for the vaccine the odds ratio is 2.178 the p value = 2.364e-12.

## Females

## Graphical

## Numerical

## Inferential

## LGBTQ

## Graphical 

## Numerical

## Inferential 

## Druggies

## Graphical

## Numerical

## Inferential 

# Overall Results and Conclusions



